---
title: "HubSpot Lead ETL"
subtitle: "Tracking Lead Status"
format: 
  html:
    theme: cosmo
    toc: true
    code-fold: true
---

Creates/updates pins for leads and lead history from HubSpot.

| object | Description | Pin URL | Full Pin URL |
|----|----|----|----|
| leads | Hubspot leads | <https://connect.procogia.com/leads/> | https://connect.procogia.com/content/51f8e6cd-0d15-419a-9a9d-7e73506dfd53 |
| lead_history | Stage changes to leads | <https://connect.procogia.com/lead_history/> | <https://connect.procogia.com/content/13b97499-4d74-47f5-90e2-deba6f136e47> |

: Pin outputs

```{r setup, message=FALSE, warning=FALSE}
library(httr2)
library(jsonlite)
library(dplyr)
library(tibble)
library(knitr)
library(purrr)
library(lubridate)
library(reactable)
library(pins)
library(gt)
library(nanoparquet)
library(tictoc)

# Load helper functions
source("R/hubspot-helpers.R")

# Set up token
hubspot_token <- Sys.getenv("HUBSPOT_ACCESS_TOKEN")
if (hubspot_token == "") {
  stop(
    "HubSpot access token not found. Please set HUBSPOT_ACCESS_TOKEN environment variable."
  )
}
tic("Total ETL time")
```

## ETL Process for HubSpot leads

### Load existing data if available

**Lookup date artificially set to Jan 1, 2025 for dev purposes.**

```{r}
tic("Load existing data")
# Check if data exists as a pin
board <- board_connect()

leads_pin_exists <- FALSE
lead_history_pin_exists <- FALSE

try(
  {
    leads <- board |> pin_read("leads") 
    leads_pin_exists <- TRUE
  },
  silent = TRUE
)

try(
  {
    lead_history <- board |> pin_read("lead_history") 
    lead_history_pin_exists <- TRUE
  },
  silent = TRUE
)

if (leads_pin_exists & lead_history_pin_exists) {
  last_lead_update <- max(leads$modified_date, na.rm = TRUE)
  cat("Loaded existing data from pins\n")
  cat("leads:", nrow(leads), "\n")
  cat("lead history records:", nrow(lead_history), "\n")
  cat("Last lead modified date:", last_lead_update, "\n")
} else {
  cat("No existing pins found, will fetch fresh data\n")
  last_lead_update <- ymd('2025-01-01')
}
toc()
```

**Limit to 200 leads for dev purposes.**

### Get the stages for leads and all leads with creation time

```{r get-leads}
# Get all leads with  time stats
all_leads <- get_all_leads_with_creation(page_limit = 100, max_pages = 2)

# Show summary
cat("Total leads:", nrow(all_leads), "\n")
cat("Ignoring leads modified before:", last_lead_update, "\n")

latest_leads <- all_leads |>
  filter(modified_date >= last_lead_update)

cat("Remaining leads:", nrow(latest_leads), "\n")
cat(
  "Date range:",
  paste0(range(latest_leads$modified_date), collapse = ' to '),
  "\n"
)


```

### Get lead histories and join with lead info

```{r lead-history}
tic("Get lead histories")
if (exists("leads") & exists("lead_history")) {
  previous_lead_data <- leads |> filter(!lead_id %in% latest_leads$lead_id)
  previous_lead_history <- lead_history |>
    filter(!lead_id %in% latest_leads$lead_id)
}
update_lead_history <- get_obj_history(latest_leads$lead_id, obj_type = "leads") |> 
  rename(lead_id = id)
update_lead_history_count <- update_lead_history |>
  count(lead_id, name = "n_history")

# Join with lead info
update_leads <- latest_leads |>
  left_join(update_lead_history_count, by = "lead_id") 

if (exists("previous_lead_data")) {
  leads <- bind_rows(previous_lead_data, update_leads)
  lead_history <- bind_rows(previous_lead_history, update_lead_history)
} else {
  leads <- update_leads
  lead_history <- update_lead_history
}
toc()
```

### Pin the data for future use

```{r pin-leads}
board <- board_connect()
board |>
  pin_write(
    leads,
    "leads",
    type = "parquet",
    description = "All leads with history count"
  )
board |>
  pin_write(
    lead_history,
    "lead_history",
    type = "parquet",
    description = "Lead history records"
  )
```

## Check if any leads don't have history

```{r}
leads |> 
  mutate(has_history = ! is.na(n_history)) |>
  count(has_history) |>
  gt()
```

## Leads Lookup

```{r}
leads |>
  reactable(
    searchable = TRUE,
    filterable = TRUE,
    columns = list(
      lead_id = colDef(name = "Lead ID"),
      lead_name = colDef(name = "Lead Name"),
      lead_type = colDef(name = "Lead Type"),
      lead_label = colDef(name = "Lead Label"),
      stage_label = colDef(name = "Stage"),
      stage_order = colDef(name = "Stage Order"),
      created_date = colDef(name = "Created"),
      modified_date = colDef(name = "Modified"),
      archived = colDef(name = "Archived"),
      n_history = colDef(name = "History Count")
    )
  )
```

## Lead History Lookup

```{r}
#default the leadID to the first lead with history
default_lead_id <- leads |> filter(n_history > 0) |> slice(1) |> pull(lead_id)

lead_history |>
  select(-sourceId) |> 
  reactable(
    searchable = TRUE,
    filterable = TRUE,
    defaultPageSize = 25,
    columns = list(
      lead_id = colDef(
        name = "lead ID",
        searchable = TRUE,
        filterable = TRUE,
      ),
      ts = colDef(name = "Timestamp"),
      stage_label = colDef(name = "Stage"),
      stage_order = colDef(name = "Stage Order"),
      source = colDef(name = "Source")
    )
  )
```

## Final timing

```{r}
toc()
```